---
title: "index"
author: "Zhe Li"
date: "2/14/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
  library(leaflet)
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(DT)
  library(scales)
```

# Reformat the data for plotting

Alaska Department of Fish and Game, Division of Commercial Fisheries. 2017. Daily salmon escapement counts from the OceanAK database, Alaska, 1921-2017. Knowledge Network for Biocomplexity. doi:10.5063/F1S46Q6M. 

```{r}
esc <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", method = "libcurl"),
stringsAsFactors = FALSE)
```

* calculate annual escapment by species and region
    + `separate`
    + `group_by`
    + `summarize` using the `sum` function 
* filter for just pacific salmon (Sockeye, Chinook, Coho, Pink, Chum)

annual_esc 

```{r}
esc_sep <- esc %>%
  separate(sampleDate, into = c("year","month","day"), sep = "-", remove = F)

head(esc_sep)
```

```{r}
annual_esc <- esc_sep %>%
  # coerce year to numeric type
  mutate(year = as.numeric(year)) %>%
  group_by(year,Species,SASAP.Region) %>%
  summarize(count = sum(DailyCount)) %>%
  filter(Species == "Sockeye" | Species == "Chinook" | Species == "Coho" | Species == "Pink" | Species == "Chum") 

head(annual_esc)
```

# Generate some static plots 

```{r}
ggplot(annual_esc, aes(x = Species, y = count, fill = SASAP.Region))+
  scale_fill_brewer(palette = "Dark2")+
  geom_col()
# fill will change the whole block 
# khaki gray8 <- color name 
# fill = "gray8"
```

```{r}
my_theme <- theme_test() +
  theme(legend.position = "bottom", 
        legend.title = element_blank(),
        plot.title = element_text(hjust = 0.5))
```


```{r}
ggplot(filter(annual_esc, SASAP.Region == "Kodiak"), aes(x=year, y=count, color=Species))+
  geom_line()+
  geom_point()+
  # the y coordinates scale, and you also can choose "percent"
  scale_y_continuous(labels = comma) + 
  ylab("Escapement (num of fish)")+
  xlab("year")+
  ggtitle("Kodiak Salmon Escapement")+
  my_theme
# theme_test()
# theme(legend.position = "bottom", legend.title = element_blank())
# theme_bw()
# theme_dark()
  
```

```{r,fig.width=5,fig.height=8}
p <- ggplot(annual_esc, aes(x=year, y=count, color=Species))+
  geom_line()+
  geom_point()+
  scale_y_continuous(labels = comma)+ 
  facet_wrap(~SASAP.Region, scales = "free_y", ncol = 2)+ 
  my_theme
p

ggsave("figures/regional_escapement.png",
       plot = p,
       width = 7,
       height = 8,
       units = "in")
```


# Generate an interactive map

